---
title: "WGCNA"
author: "Bridget Bai"
date: "3/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#libraries
```{r}
library(gplots)
library(ggplot2)
library(WGCNA)
options(stringsAsFactors = FALSE)
library(AnnotationDbi)
library(topGO)
library(viridis)
```

####prepare data for analysis

#import protein data 
```{r}
df_WGCNA <- read.csv("~/Desktop/rotation_Briggs/groups_organized.csv", header = TRUE, stringsAsFactors = FALSE)
df_sub <- df_WGCNA[,c(1,4:18)]

df_sub <- df_sub[!df_sub$B317 == "#DIV/0!", ]
df_sub <- df_sub[!df_sub$B21 == "#DIV/0!", ]
df_sub <- df_sub[!df_sub$B14 == "#DIV/0!", ]
df_sub <- df_sub[!df_sub$BM_6 == "#DIV/0!", ]
df_sub <- df_sub[!df_sub$B16 == "#DIV/0!", ]
df_sub <- df_sub[!df_sub$BM == "#DIV/0!", ]

rownames(df_sub) <- df_sub$Accession
df_sub[,2:16] <- lapply(df_sub[,2:16], as.numeric)
df_sub <- df_sub[, -1]
df_sub <- data.frame(t(df_sub))
#df_sub[,1:3920] <- lapply(df_sub[,1:3920], as.numeric)
```
#import traits file 
```{r}
hyb_traits <- read.csv("HMP_traits.csv", header = TRUE, stringsAsFactors = FALSE)
rownames(hyb_traits) <- hyb_traits$sample

barplot(hyb_traits$plant.height,space = 0)
```

#pick soft-thresholding powers
```{r}
WGCNA_plot1=function(df_sub){
    # Choose a set of soft-thresholding powers
    powers = c(c(4:16), seq(from = 18, to=25, by=2))
    # Call the network topology analysis function
    sft = WGCNA::pickSoftThreshold(df_sub, powerVector = powers, verbose = 5)

    #plot
    require(repr)
    tiff('pick_beta_power.tiff', units = "in", width = 8, height = 6, res = 300, compression = "lzw")
    options(repr.plot.width=9, repr.plot.height=4)
    par(mfrow = c(1,2));
    cex1 = 0.9;
    plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
    xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
    main = paste("Scale independence"));
    text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
    labels=powers,cex=cex1,col="red");
    # this line corresponds to using an R^2 cut-off of h
    abline(h=0.89,col="red")
    # Mean connectivity as a function of the soft-thresholding power
    plot(sft$fitIndices[,1], sft$fitIndices[,5],
    xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
    main = paste("Mean connectivity"))
    text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
    dev.off()
}

WGCNA_plot1(df_sub)
#softpower = 12
```

####generate network and module

#calculate adjacency 
```{r}
softPower = 12 #see figure in last chunk
adjacency <-WGCNA::adjacency(df_sub, power = softPower, type = "signed")
```

#topological overlap matrix
```{r}
#turn adjacency into topological overlap
dissTOM <-  TOMdist(adjacency)
```

#clustering using TOM
```{r}
# hierarchical clustering
geneTree = hclust(as.dist(dissTOM), method = "average")
# define the modules by cutting branches
moduleLabelsManual1 = cutreeDynamic(dendro = geneTree, distM = dissTOM, method = "hybrid", 
    deepSplit = 2, pamRespectsDendro = F, minClusterSize = 30)
moduleColorsManual1 = labels2colors(moduleLabelsManual1)
```

#calculate eigengenes and merge modules based on similar expression profiles
```{r}
# Calculate eigengenes
MEList <- moduleEigengenes(df_sub, colors = moduleColorsManual1)
MEs <- MEList$eigengenes

#cut tree to merge modules
MEDissThres = 0.15
# Call an automatic merging function
merge = mergeCloseModules(df_sub, moduleColorsManual1, cutHeight = MEDissThres)
# The merged module colors
mergedColors = merge$colors
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs
```

#quantifying module-trait association
```{r}
moduleColors = mergedColors

# Define numbers of genes and samples
nGenes = ncol(df_sub)
nSamples = nrow(df_sub)
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(df_sub, moduleColors)$eigengenes
MEs = orderMEs(MEs0)

moduleTraitCor = cor(MEs,hyb_traits$plant.height ,use = "p")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)

#summarize correlation and p-value for all modules
moduleTrait_summary <- cbind(moduleTraitCor, moduleTraitPvalue)
colnames(moduleTrait_summary) <- c("Correlation_height","P-value")
write.csv(moduleTrait_summary, "moduleTrait_summary.csv")
```

#module-trait heatmap
````{r}

MET = orderMEs(cbind(mergedMEs[, c("MEturquoise","MEred","MEyellow","MEbrown","MEgreen")], hyb_traits$plant.height))
colnames(MET)[4] <- "plant height"
MET_matrix <- cor(MET,MET)     
MET_matrix_pvalue <- corPvalueStudent(MET_matrix, nSamples)

textMatrix = paste(signif(MET_matrix, 2), "\n(",
                        signif(MET_matrix_pvalue, 1), ")", sep = "")  

tiff('module_height_heatmap.tiff', units = "in", width = 10, height = 6, res = 300, compression = "lzw")
par(mar = c(6, 7, 2, 1))
labeledHeatmap(met_matrix,xLabels = c("MEturquoise","MEred","MEyellow","plant height","MEbrown","MEgreen"),yLabels = c("MEturquoise","MEred","MEyellow","plant height","MEbrown","MEgreen"),colorLabels = FALSE,
             colors = viridis(30,alpha = 0.8,begin = 0.4,end = 1),
             textMatrix = textMatrix,
             setStdMargins = FALSE,
             cex.text = 1,
             ySymbols = c("404","214","295","","314","246"),xSymbols =c("404","214","295","","314","246"))
```

# calculate the module membership values (aka. module eigengene based connectivity kME) & Gene Significance
```{r}
datKME = signedKME(df_sub, MEs)
modNames = substring(names(datKME), 4)
G.S_height = as.numeric(cor(df_sub, hyb_traits$plant.height, use = "p"))

par(mfrow = c(2, 2))
selectModules = c("yellow", "red", "turquoise", "brown", "green")
for (module in selectModules) {
    column = match(module, modNames)
    restModule = moduleColors == module
    tiff(paste0("MM_vs_GS",module,".tiff"), units = "in", width = 10, height = 6, res = 300, compression = "lzw")
    verboseScatterplot(datKME[restModule, column], G.S_height[restModule], xlab = paste("Module Membership ", 
        module, "module"), ylab = "GS.plant.height", main = paste("kME.", module, 
        "vs. GS"), col = module,cex.lab = 1,cex.main = 0.8,cex.axis = 1)
    dev.off()
}
```

#write module data to file
```{r}
probes = colnames(df_sub)
probes2annot = match(probes, df_WGCNA$Accession)
sum(is.na(probes2annot))
datGS.Traits = data.frame(cor(df_sub, hyb_traits$plant.height, use = "p"))
colnames(datGS.Traits) <- "geneTraitSignificance"
datOutput = data.frame(ProbeID = probes, moduleColors, datKME, datGS.Traits)
write.csv(datOutput,file = "WGCNA_leaf_summary.csv")
```

#read in protein annotation files
```{r}
SL_GOI <- read.csv(file = "~/Desktop/rotation_Briggs/FinalLeafGOIs.csv", header = TRUE, stringsAsFactors = FALSE)
```

#merge individual module with annotation
```{r}
merge_module_GO <- function(a,b){
  df_1 <- datOutput[datOutput$moduleColors == a, c("ProbeID","moduleColors","geneTraitSignificance",b)]
  df_2 <- merge.data.frame(x = df_1, y = SL_GOI[, c("Accession","Protein","Category","GOterm","KeggPathway")], by.x = "ProbeID", by.y  = "Accession")  
  df_2 <- unique(df_2)
  df_2 <- df_2[order(df_2[,b],decreasing = TRUE), ]
  df_2$Gene <- substr(df_2$ProbeID, start = 1, stop = 14)
  df_2 <- df_2[-(grep("only",df_2[,"ProbeID"])), ]
  df_2
}

yellow <- merge_module_GO(a = "yellow",b = "kMEyellow")
brown <- merge_module_GO(a = "brown", b = "kMEbrown")
turquoise <- merge_module_GO(a = "turquoise", b = "kMEturquoise")
green <- merge_module_GO(a = "green", b = "kMEgreen")
red <- merge_module_GO(a = "red", b = "kMEred")


```



#separate top modules 
```{r}
yellow <- merge_module_GO(a = "yellow",b = "kMEyellow")
brown <- merge_module_GO(a = "brown", b = "kMEbrown")
turquoise <- merge_module_GO(a = "turquoise", b = "kMEturquoise")
green <- merge_module_GO(a = "green", b = "kMEgreen")
red <- merge_module_GO(a = "red", b = "kMEred")

#hub_genes <- data.frame(rbind(brown,turquoise,red,yellow,green))
hub_genes <- data.frame(rbind(hub_genes,blue))
#write.csv(hub_genes, file = "hub_genes.csv")
module_gene <- data.frame(rbind(yellow,brown,turquoise,green,red))
write.csv(module_gene,file = "module_gene.csv")

#mean(green$geneTraitSignificance)
```

#load gamer
```{r}
GAMER <- read.delim("~/Desktop/rotation_Briggs/Mt_analysis/maize.B73.AGPv4.aggregate.gaf", sep = "\t", header = FALSE, stringsAsFactors = F)
colnames(GAMER) <- c("db", "db_object_id", "db_object_symbol", "qualifier", "term_accession", "db_reference", "evidence_code", "with", "aspect", "db_object_name", "db_object_synonym", "db_object_type", "taxon", "date", "assigned_by", "annotation_extension", "gene_product_form_id")
gamer <- GAMER[-(1:2),]
gamer$Term <- Term(as.character(gamer$term_accession))
gamercut <- gamer[gamer$db_object_id %in% df_sub$gene, ]
```
#GO function
```{r}
GOanalysisGodata <- function(EnrichedDF, GamerCut, ont, nodesize = 1){
  gene2GO_TissueCut <- split((as.character(GamerCut$term_accession)),GamerCut$db_object_id)
  gene2GOb_TissueCut <- gene2GO_TissueCut[-2]
  str(head(gene2GOb_TissueCut))
  geneNames_tissuecut <- names(gene2GOb_TissueCut)
  myInterestingGenes <- EnrichedDF$Gene
  geneList <- factor(as.integer(geneNames_tissuecut %in% myInterestingGenes))
  names(geneList) <- geneNames_tissuecut
  GOdata <- new("topGOdata", ontology = ont, allGenes = geneList, annot = annFUN.gene2GO, nodeSize=nodesize, gene2GO = gene2GOb_TissueCut)
  GOdata
}


```


#GO term enrichment 
```{r}
#yellow
yellow_go_bp <- GOanalysisGodata(yellow, gamercut, "BP", nodesize = 5)
yellow_go_cc <- GOanalysisGodata(yellow, gamercut, "CC", nodesize = 5)
yellow_go_mf <- GOanalysisGodata(yellow, gamercut, "MF", nodesize = 5)
yellow_bp_stats <- GOanalysisDF(yellow_go_bp,"yellow")
yellow_cc_stats <- GOanalysisDF(yellow_go_cc,"yellow")
yellow_mf_stats <- GOanalysisDF(yellow_go_mf,"yellow")
yellow_go_stats <- rbind(yellow_bp_stats, yellow_cc_stats, yellow_mf_stats)


#green
green_go_bp <- GOanalysisGodata(green, gamercut, "BP", nodesize = 5)
green_go_cc <- GOanalysisGodata(green, gamercut, "CC", nodesize = 5)
green_go_mf <- GOanalysisGodata(green, gamercut, "MF", nodesize = 5)
green_bp_stats <- GOanalysisDF(green_go_bp, "green")
green_cc_stats <- GOanalysisDF(green_go_cc, "green")
green_mf_stats <- GOanalysisDF(green_go_mf, "green")
green_go_stats <- rbind(green_bp_stats, green_cc_stats, green_mf_stats)


#turquoise
turquoise_go_bp <- GOanalysisGodata(turquoise, gamercut, "BP", nodesize = 5)
turquoise_go_cc <- GOanalysisGodata(turquoise, gamercut, "CC", nodesize = 5)
turquoise_go_mf <- GOanalysisGodata(turquoise, gamercut, "MF", nodesize = 5)
turquoise_bp_stats <- GOanalysisDF(turquoise_go_bp,"turquoise")
turquoise_cc_stats <- GOanalysisDF(turquoise_go_cc,"turquoise")
turquoise_mf_stats <- GOanalysisDF(turquoise_go_mf,"turquoise")
turquoise_go_stats <- rbind(turquoise_bp_stats, turquoise_cc_stats, turquoise_mf_stats)


red_go_bp <- GOanalysisGodata(red, gamercut, "BP", nodesize = 5)
red_go_cc <- GOanalysisGodata(red, gamercut, "CC", nodesize = 5)
red_go_mf <- GOanalysisGodata(red, gamercut, "MF", nodesize = 5)
red_bp_stats <- GOanalysisDF(red_go_bp,"red")
red_cc_stats <- GOanalysisDF(red_go_cc,"red")
red_mf_stats <- GOanalysisDF(red_go_mf,"red")
red_go_stats <- rbind(red_bp_stats, red_cc_stats, red_mf_stats)

#brown
brown_go_bp <- GOanalysisGodata(brown, gamercut, "BP", nodesize = 5)
brown_go_cc <- GOanalysisGodata(brown, gamercut, "CC", nodesize = 5)
brown_go_mf <- GOanalysisGodata(brown, gamercut, "MF", nodesize = 5)
brown_bp_stats <- GOanalysisDF(brown_go_bp,"brown")
brown_cc_stats <- GOanalysisDF(brown_go_cc,"brown")
brown_mf_stats <- GOanalysisDF(brown_go_mf,"brown")
brown_go_stats <- rbind(brown_bp_stats, brown_cc_stats, brown_mf_stats)

GO_Summary_WGCNA <- rbind(yellow_go_stats, green_go_stats, turquoise_go_stats, red_go_stats, brown_go_stats)
write.csv(GO_Summary_WGCNA, file = "GO_Summary_WGCNA.csv")
```

#find hub genes
```{r}
find_hub_gene <- function(df, modulekME){
  df <- df[!(abs(df[,modulekME]) <= 0.9), ]
  df <- df[!(abs(df[,"geneTraitSignificance"]) <= 0.5), ]
  colnames(df)[4] <- "module membership"
  df
}

yellow_hub <- find_hub_gene(yellow, "kMEyellow")
red_hub <- find_hub_gene(red, "kMEred")
green_hub <- find_hub_gene(green, "kMEgreen")
brown_hub <- find_hub_gene(brown, "kMEbrown")
turquoise_hub <- find_hub_gene(turquoise, "kMEturquoise")

hub_genes <- data.frame(rbind(yellow_hub, red_hub, green_hub, brown_hub, turquoise_hub))
write.csv(hub_genes, "hub_genes_WGCNA.csv")  
```

#export to cytoscape
```{r}
modules = c("yellow", "red","turquoise","brown","green")
inModule = is.finite(match(moduleColors, modules))
modProbes = probes[inModule]
modTOM = adjacency[inModule, inModule]
dimnames(modTOM) = list(modProbes, modProbes)
cyt = exportNetworkToCytoscape(modTOM, weighted = TRUE, nodeAttr = moduleColors[inModule])
node_data <- cyt$nodeData
edge_data <- cyt$edgeData
write.csv(edge_data,"cytoscape_edge_input.csv")
write.csv(node_data,"cytoscape_node_input.csv")
```

#heatmap to compare between hybrids
```{r}
df_sub_1 <- df_sub
rownames(df_sub_1) <- c("BM","MB","B84B","B84M","A682B","A682M","RIL_BM","14B","14M","16B","16M","21B","21M","317B","317M")
rownames(df_sub_1)[1:6] <- paste0("6hybrid_",rownames(df_sub_1)[1:6])
rownames(df_sub_1)[8:15] <- paste0("RIL_",rownames(df_sub_1)[8:15])

adjust_range <- function(a){
  if (a < 0.9){
    a = 0.9
  }else if
   (a > 1.2){
    a = 1.2
   }
  a
}

#hub gene heatmap
hub_gene_select <- hub_genes$ProbeID
df_sub_hub <- df_sub_1[,hub_gene_select]
hub_colors <- hub_genes$moduleColors
df_sub_hub_adjusted <- apply(df_sub_hub,c(1,2),adjust_range)
df_sub_interest_hub_ordered <- df_sub_hub_adjusted[c(9,10,11,13,15,7,3,6,1,2,4,5,8,12,14),]


colors = seq(0.9,1.2,length = 101)
my_palette <- colorRampPalette(c("yellow","gray","darkblue"))(n = 100)
par(mar = c(5,25,3,10))
heatmap.2(df_sub_hub_adjusted_ordered, Colv = T,reorderfun=function(d,w) reorder(d,w,agglo.FUN = mean),rowsep = c(4,12),dendrogram = "both",scale = "none", trace = "none",ColSideColors = hub_colors, density.info = "none", col = my_palette, breaks = colors,cexRow = 1.2,offsetCol = 0,offsetRow = -0.4, margins = c(7,7.5), labCol = NA, keysize = 1.5,key.xlab = "HYB/MP",key.xtickfun=function() {
               cex <- par("cex")*par("cex.axis")
               side <- 1
               line <- 0
               col <- par("col.axis")
               font <- par("font.axis")
               mtext("<0.9", side=side, at=0, adj=0,
                     line=line, cex=cex, col=col, font=font)
              mtext("1", side=side, at=0.5, adj=0.5,
                     line=line, cex=cex, col=col, font=font)
               mtext(">1.2", side=side, at=1, adj=1,
                     line=line, cex=cex, col=col, font=font)
               return(list(labels=FALSE, tick=FALSE))
          })

#all gene heatmap

df_sub_all_gene <- df_sub_1[,inModule]
df_sub_all_gene <- df_sub_all_gene[,-grep("only",colnames(df_sub_all_gene))]
select_colors <- moduleColors[inModule]
select_colors <- select_colors[-grep("only",modProbes)]
df_sub_all_adjusted <- apply(df_sub_all_gene,c(1,2),adjust_range)
df_sub_all_ordered <- df_sub_all_adjusted[c(9,10,11,13,15,7,3,6,1,2,4,5,8,12,14),]

par(mar = c(5,25,3,10))
heatmap.2(df_sub_all_ordered, Colv = T,reorderfun=function(d,w) reorder(d,w,agglo.FUN = mean),rowsep = c(4,12),dendrogram = "both",scale = "none", trace = "none",ColSideColors = select_colors, density.info = "none", col = my_palette, breaks = colors,cexRow = 1.2,offsetCol = 0,offsetRow = -0.4, margins = c(7,7.5), labCol = NA, keysize = 1.5,key.xlab = "HYB/MP",key.xtickfun=function() {
               cex <- par("cex")*par("cex.axis")
               side <- 1
               line <- 0
               col <- par("col.axis")
               font <- par("font.axis")
               mtext("<0.9", side=side, at=0, adj=0,
                     line=line, cex=cex, col=col, font=font)
              mtext("1", side=side, at=0.5, adj=0.5,
                     line=line, cex=cex, col=col, font=font)
               mtext(">1.2", side=side, at=1, adj=1,
                     line=line, cex=cex, col=col, font=font)
               return(list(labels=FALSE, tick=FALSE))
          })





#barplots to merge with heatmap
hyb_traits_reordered <- hyb_traits[c(5,4,2,1,15,11,9,10,13,3,6,7,12,14,8),]
par(mar = c(3,5,3,2))
barplot(hyb_traits_reordered$plant.height, ylab = "HYB/MP Plant Height",
col = c("black","black","black","black","gray","gray","gray","gray","gray","gray","gray","gray","white","white","white"), font.lab = 2, cex.axis = 1.6, cex.lab = 2)

```

